---
title: "Fitting model to data"
author: "anjie"
date: "8/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(matrixStats)
library(partitions)
library(tictoc)
library(LaplacesDemon)

source(here("helper/make_scheme_and_params.R"))
source(here("helper/initialization.R"))
source(here("helper/probability_computations.R"))
source(here("helper/main_simulation_helper.R"))

d <- read_csv(here("data/processed_rt_task_data.csv"))

# set up multicore sims
future::plan(multicore)
```

Figuring out the sequence and generate the stimuli sequence with all the parameters 

```{r}

# set model-related parameters 
alpha_epsilons = c(1)
beta_epsilons = c(10)
alpha_priors = c(1)
beta_priors = c(30)
noise_parameters = c(0)
world_EIGs = c(0.001)
max_observation = 100

model_params <- set_model_params(alpha_priors, beta_priors, 
                alpha_epsilons, beta_epsilons, 
                noise_parameters, world_EIGs, max_observation)


# set stimuli-related parameters 
features_df <- tibble(
  n_features = c(5, 5), 
  on_features_n = c(1, 3)
)
sequence_scheme = c("BDBBBB", "BBBDBB", "BBBBBD", "BBBBBB")

stims_df <- set_stim_params(sequence_scheme, features_df)


full_params_df <- make_simulation_params(n_sim = 150,
                                        model_params, 
                                        stims_df)

# option 1: make sure the simple and the complex one has the same number of flipped feature 
  # should reinforce in the feature_df section to make sure simulations are relatively independent of each other? 

# option 2: having all features flipped, when plausible. then poses a limitation on the number of features turned on (n_on <= n_total * .5)
  # currently following the option 2. if it is above n_total * .5 then flip the maximum of the feature 



```



Running the simulation. 

```{r}
tic()
all_sims_res <- full_params_df %>%
  mutate(row_number = row_number()) %>% 
  group_by(row_number) %>% 
  nest() %>%
  mutate(results = map(data,
                       function(df) main_simulation(params = df))) %>%
  unnest(cols = c(data, results))
toc()
```

Clean and visualize. 

```{r}
all_sims_res %>% 
  group_by(row_number, params_id, n_features, on_features_n, sequence_scheme,  stimulus_idx) %>% 
  filter(!is.na(stimulus_idx)) %>% 
  summarise(sample_n = n()) %>% 
  ggplot(aes(x = stimulus_idx, y = sample_n, group = as.factor(on_features_n))) + 
  stat_summary(aes(color = as.factor(on_features_n)), position =position_dodge(width = .1),  fun.data = "mean_cl_boot") + 
  facet_wrap(~sequence_scheme)


  #geom_line(alpha = .1, aes(color = n_features)) + 
```
observation: 
- dishabituation is still very weak 
- weird tail going up that doesn't make sense 
- complexity showing up in the dishabituation trial (kind of) but not really 

first look into the weird tail going up situation
```{r}

# set model-related parameters 
alpha_epsilons = c(1)
beta_epsilons = c(10)
alpha_priors = c(1)
beta_priors = c(30)
noise_parameters = c(0)
world_EIGs = c(0.001)
max_observation = 100

model_params <- set_model_params(alpha_priors, beta_priors, 
                alpha_epsilons, beta_epsilons, 
                noise_parameters, world_EIGs, max_observation)


# set stimuli-related parameters 
features_df <- tibble(
  n_features = c(3), 
  on_features_n = c(1)
)
sequence_scheme = c("BB", "BBB", "BBBBB", "BBBBBB", "BBBBBBB", "BBBBBBBB", "BBBBBBBB")

stims_df <- set_stim_params(sequence_scheme, features_df)


tail_params_df <- make_simulation_params(n_sim = 150,
                                        model_params, 
                                        stims_df)


```


```{r}
tail_sims_res <- tail_params_df %>%
  mutate(row_number = row_number()) %>% 
  group_by(row_number) %>% 
  nest() %>%
  mutate(results = map(data,
                       function(df) main_simulation(params = df))) %>%
  unnest(cols = c(data, results))
```

```{r}
tail_sims_res %>% 
  group_by(row_number, params_id, n_features, on_features_n, sequence_scheme,  stimulus_idx) %>% 
  filter(!is.na(stimulus_idx)) %>% 
  summarise(sample_n = n()) %>% 
  ggplot(aes(x = stimulus_idx, y = sample_n, group = as.factor(on_features_n))) + 
  stat_summary(aes(color = as.factor(on_features_n)), position =position_dodge(width = .1),  fun.data = "mean_cl_boot") + 
  facet_wrap(~sequence_scheme)
```

```{r}

# set model-related parameters 
alpha_epsilons = c(1)
beta_epsilons = c(10)
alpha_priors = c(1)
beta_priors = c(15)
noise_parameters = c(0)
world_EIGs = c(0.001)
max_observation = 100

model_params <- set_model_params(alpha_priors, beta_priors, 
                alpha_epsilons, beta_epsilons, 
                noise_parameters, world_EIGs, max_observation)


# set stimuli-related parameters 
features_df <- tibble(
  n_features = c(3), 
  on_features_n = c(1)
)
sequence_scheme = c("BB", "BBB", "BBBBB", "BBBBBB", "BBBBBBB", "BBBBBBBB", "BBBBBBBB")

stims_df <- set_stim_params(sequence_scheme, features_df)


tail_params_df <- make_simulation_params(n_sim = 150,
                                        model_params, 
                                        stims_df)


```


```{r}
tail_sims_res_a1b15 <- tail_params_df %>%
  mutate(row_number = row_number()) %>% 
  group_by(row_number) %>% 
  nest() %>%
  mutate(results = map(data,
                       function(df) main_simulation(params = df))) %>%
  unnest(cols = c(data, results))
```

```{r}
tail_sims_res_a1b15 %>% 
  group_by(row_number, params_id, n_features, on_features_n, sequence_scheme,  stimulus_idx) %>% 
  filter(!is.na(stimulus_idx)) %>% 
  summarise(sample_n = n()) %>% 
  ggplot(aes(x = stimulus_idx, y = sample_n, group = as.factor(on_features_n))) + 
  stat_summary(aes(color = as.factor(on_features_n)), position =position_dodge(width = .1),  fun.data = "mean_cl_boot") + 
  facet_wrap(~sequence_scheme)
```



```{r}

# set model-related parameters 
alpha_epsilons = c(1)
beta_epsilons = c(10)
alpha_priors = c(1, 10, 30)
beta_priors = c(1, 10, 30)
noise_parameters = c(0)
world_EIGs = c(0.001)
max_observation = 100

model_params <- set_model_params(alpha_priors, beta_priors, 
                alpha_epsilons, beta_epsilons, 
                noise_parameters, world_EIGs, max_observation)


# set stimuli-related parameters 
features_df <- tibble(
  n_features = c(1, 3), 
  on_features_n = c(1, 3)
)
sequence_scheme = c("BDBBBB")

stims_df <- set_stim_params(sequence_scheme, features_df)


full_params_df <- make_simulation_params(n_sim = 150,
                                        model_params, 
                                        stims_df)


```



Running the simulation. 

```{r}
tic()
all_sims_res <- full_params_df %>%
  mutate(row_number = row_number()) %>% 
  group_by(row_number) %>% 
  nest() %>%
  mutate(results = map(data,
                       function(df) main_simulation(params = df))) %>%
  unnest(cols = c(data, results))
toc()
```

Clean and visualize. 

```{r}
all_sims_res %>% 
  mutate(params_info = paste0("a", alpha_prior, "b", beta_prior)) %>% 
  group_by(row_number, params_id,params_info, n_features, sequence_scheme,  stimulus_idx) %>% 
  filter(!is.na(stimulus_idx)) %>% 
  summarise(sample_n = n()) %>% 
  ggplot(aes(x = stimulus_idx, y = sample_n, group = as.factor(n_features))) + 
  stat_summary(aes(color = as.factor(n_features)), position =position_dodge(width = .1),  fun.data = "mean_cl_boot") + 
  facet_wrap(sequence_scheme~params_info)


  #geom_line(alpha = .1, aes(color = n_features)) + 
```






