---
title: "find good grid"
author: "anjie"
date: '2022-06-22'
output: html_document
---

```{r}
library(here)
library(tictoc)
library(tidyverse)
library(matrixStats)
source(here("helper/compute_prob.r"))
source(here("helper/init.r"))
source(here("helper/granch_main_simulation_helper.R")) # this will import other helper functions 
source(here("helper/granch_sim_setup_helper.R")) # this will import functions that help with setting parameters 

source(here("helper/debugging/format_helper.r"))
source(here("helper/debugging/visualization_helper.r"))

```

# Set up stimuli pool

```{r}
embedding_path <- here("embeddings/embedding_PCA.csv")
stimuli_pool_size <- 10
hypothetical_obs_grid_n <- 3
n_feature <- 1
stimuli_pool <-  get_stimuli_pool(stimuli_pool_size, n_feature, embedding_path)

stimuli_pool <- scale(stimuli_pool, center = TRUE, scale = colSums(stimuli_pool))
  
  
# sd_n: range decided by n*sd away 
# sd_step: step decided by sd / sd_step 
grid_mu_theta = get_grid_mu_theta(stimuli_pool, sd_n = 5, sd_step = 3)

# not sure about the principled way to select these three yet
grid_sig_sq = seq(0.02, 1, 0.02) # not sure about the principled way to select this yet, but anything too small, like below 0.02 will give-f
grid_y <- grid_mu_theta 
# grid_epsilon might be too small 
grid_epsilon = seq(0.0001, 1, 0.2)

# set model-related parameters 
em <- read_csv(embedding_path, col_names = FALSE)
mean_em <- mean(as.matrix(em))

mu_priors = c(mean_em)
mu_priors = c(0)
V_priors = c(0.5) # maybe this controls for how fast?
alpha_priors = c(2) 
beta_priors = c(2) 
epsilons = c(0.0001) 
mu_epsilon = c(0)
sd_epsilon = c(.5)
world_EIGs = c(1) 
max_observation = 500

model_params <- set_granch_params(
                grid_mu_theta, grid_sig_sq, grid_y, grid_epsilon, hypothetical_obs_grid_n,
                mu_priors, V_priors, 
                alpha_priors, beta_priors, 
                epsilons, mu_epsilon, sd_epsilon,
                world_EIGs, max_observation)

model_params <- add_precalculated_prior_dfs(model_params)

```

```{r}

b <- 0.1
d <- 0.3

stims_df <- tibble(sequence_scheme = c("BBB"),
                   n_features = n_feature
                   ) %>% 
  mutate(
    stimuli_sequence = map(sequence_scheme, function(ss){make_real_stimuli_df(ss,background = b  ,
deviant =d
)}))



full_params_df <- make_simulation_params(n_sim = 1,
                                        model_params, 
                                        stims_df)
```

```{r}
full_params_df$stimuli_sequence

```

# backward im metrics
}else if(measurement == "KL"){
    model$KL = rep(NA,max_observation)
  }else if(measurement == "surprisal"){
    model$surprisal = rep(NA,max_observation)

```{r}

full_params_df$im <- "surprisal"

```




## density

```{r}


tibble(grid_mu_theta = grid_mu_theta,
line  = 0) %>%
  ggplot(aes(x = grid_mu_theta, 
             y = line)) + 
  geom_point() + 
  geom_vline(xintercept = b, color = "red"	) + 
  geom_vline(xintercept =d, color = "blue") +
  scale_x_continuous(breaks = unique(grid_mu_theta)) + 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank())+ 
  ylab("")

  

```

# backward model 

```{r}
tic()
all_sims_res <- full_params_df %>%
  mutate(row_number = row_number()) %>% 
  group_by(row_number) %>% 
  nest() %>%
  mutate(results = map(data,
                       function(df) granch_backward_im_simulation(params = df))) %>%
 unnest(cols = c(data, results))
toc()
```

```{r}
all_sims_res
```



# forward model 

```{r}
tic()
all_sims_res <- full_params_df %>%
  mutate(row_number = row_number()) %>% 
  group_by(row_number) %>% 
  nest() %>%
  mutate(results = map(data,
                       function(df) granch_main_simulation(params = df))) %>%
  unnest(cols = c(data, results))
toc()
```



```{r}
plot_sim_results(all_sims_res)
plot_sim_eig(all_sims_res)
```

current focus is to understand why doesn't first one habituate? 


```{r}
sim1_posteriors <- get_posterior(all_sims_res, 1)
sim1_hypo_posteriors <- get_hypo_posterior(all_sims_res, 1)


sim1_first_transition <- nrow(get_stimulus_idx_t(all_sims_res, 1, 1)) + 1 # gives us the first trial for the 
sim1_second_transition <- max(get_stimulus_idx_t(all_sims_res, 1, 2)$t) + 1 # gives us the first trial for the 
sim1_third_transition <- max(get_stimulus_idx_t(all_sims_res, 1, 3)$t) + 1 # gives us the first trial for the 
sim1_fourth_transition <- max(get_stimulus_idx_t(all_sims_res, 1, 4)$t) + 1 # gives us the first trial for the 


sim1_eig_componetns <- format_hypothetical_kl_pp(all_sims_res, 1)
sim1_kl <- format_hypothetical_kl_pp(all_sims_res, 1) %>% filter(type == "kl") %>% 
  pivot_longer(cols = starts_with("V"), names_to = "feature", values_to = "kl_val") 
sim1_pp <- format_hypothetical_kl_pp(all_sims_res, 1) %>% filter(type == "pp") %>% 
  pivot_longer(cols = starts_with("V"), names_to = "feature", values_to = "pp_val") 
```


```{r}


sim1_kl %>% 
  filter(timestep >= sim1_second_transition, timestep < sim1_fourth_transition) %>% 
  ggplot(aes(x = timestep, y = kl_val)) +
  geom_point()+ 
  facet_wrap(~feature)

sim1_pp %>% 
  filter(timestep >= sim1_second_transition,timestep < sim1_fourth_transition) %>% 
  ggplot(aes(x = timestep, y = pp_val)) +
  geom_point()+ 
  facet_wrap(~feature)

plot_sim_eig_on_stimulus(all_sims_res, 1, 3)

```

very weird shapes of kl on the second observation on the same stimulus, why going up and then going down? 


```{r}
visualize_posteriors(sim1_posteriors, sim1_first_transition + 2, sim1_second_transition)
visualize_hypo_posteriors(sim1_hypo_posteriors, sim1_first_transition + 2, sim1_second_transition)
```

the posteriors all look great, ok, let's now do some hand calculation

```{r}
p23 <- sim1_posteriors %>% 
  filter(timestep == sim1_first_transition + 2)
hp23 <- sim1_hypo_posteriors %>% 
  filter(timestep == sim1_first_transition + 2, hypo_obs_id ==1)


p23$posterior[p23$posterior < exp(-700)] <- 1/(10^300)
hp23$posterior[hp23$posterior < exp(-700)] <- 1/(10^300)

p24 <- sim1_posteriors %>% 
  filter(timestep == sim1_first_transition + 3)
hp24 <- sim1_hypo_posteriors %>% 
  filter(timestep == sim1_first_transition + 3, hypo_obs_id ==1)

p24$posterior[p24$posterior < exp(-700)] <- 1/(10^300)
hp24$posterior[hp24$posterior < exp(-700)] <- 1/(10^300)

        
kl_div(hp23$posterior, p23$posterior)
kl_div(hp24$posterior, p24$posterior)

```





```{r}

format_hypothetical_kl_pp(all_sims_res, 1) %>% 
  filter(type == "kl") %>% 
  filter(timestep > 30) %>% 
  pivot_longer(cols = starts_with("V"), names_to = "feature", values_to = "kl_val") %>% 
  ggplot(aes(x = timestep, y = kl_val)) +
  geom_point()+ 
  facet_wrap(~feature)

format_hypothetical_kl_pp(all_sims_res, 2) %>% 
  filter(type == "kl") %>% 
  filter(timestep > 30) %>% 
  pivot_longer(cols = starts_with("V"), names_to = "feature", values_to = "kl_val") %>% 
  ggplot(aes(x = timestep, y = kl_val)) +
  geom_point()+ 
  facet_wrap(~feature)


format_hypothetical_kl_pp(all_sims_res, 1) %>% 
  filter(type == "pp") %>% 
  filter(timestep > 30) %>% 
  pivot_longer(cols = starts_with("V"), names_to = "feature", values_to = "pp_val") %>% 
  ggplot(aes(x = timestep, y = pp_val)) +
  geom_point()+ 
  facet_wrap(~feature)

format_hypothetical_kl_pp(all_sims_res, 2) %>% 
  filter(type == "pp") %>% 
  filter(timestep > 30) %>% 
  pivot_longer(cols = starts_with("V"), names_to = "feature", values_to = "pp_val") %>% 
  ggplot(aes(x = timestep, y = pp_val)) +
  geom_point()+ 
  facet_wrap(~feature)
```





