---
title: "EIG computation"
author: "anjie & mike"
date: "6/3/2021"
output: html_document
---

# Preliminaries

```{r}
# library(reshape2)
library(tidyverse)
library(here)
library(matrixStats)
```


```{r}
## source relevant files
source(here("helper/get_entropy.R"))
source(here("helper/get_KL_measurement.R"))
source(here("helper/get_surprise.R"))
source(here("helper/noisy_update.R"))
source(here("helper/grid_approximation.R"))
source(here("helper/get_stimuli_and_observations.R"))
```

# Generate scenario

```{r}
num_features = 1
num_features_simple = 1
# num_features_complex = 8
trials_per_block = 8
deviant_positions = 4
dissimilarity_ratio = 0.2
epsilon = 0.1

# must satisfy: total feature > (1 + dissimilar ratio) * featureOnumber
simple_stimuli <- generate_creature_sequence(
  block_length = trials_per_block, 
  deviant_positions = deviant_positions,  # takes a vector, 
  total_feature = num_features, 
  feature_theta = 0.8, 
  feature_number = num_features_simple, 
  dissimilar_ratio = dissimilarity_ratio)


simple_observations <- 
  generate_noisy_observations(block = simple_stimuli, 
                              exposure_type = "self_paced", 
                              short_exposure_samps = 1, 
                              long_exposure_samps = 10, 
                              normal_exposure_samps = 10, 
                              epsilon = epsilon)    
```

# Posterior inference

First start by computing the posterior after observation 1. 

```{r}
grid_theta <- seq(0.1, 1, 0.2)
grid_epsilon <- seq(0.1, 1, 0.2)
alpha_prior = 1
beta_prior = 1
alpha_epsilon = 1 
beta_epsilon = 10

obs_matrix <- simple_observations %>% 
  ungroup() %>%
  select(-c(trial_num, observation_num)) %>% 
  as.matrix()

posterior_o1 <- grid_with_theta_and_epsilon(grid_theta = grid_theta, 
                                            grid_epsilon = grid_epsilon, 
                                            noisy_observation = obs_matrix[1,], 
                                            alpha_prior = alpha_prior, 
                                            beta_prior= beta_prior, 
                                            alpha_epsilon = alpha_epsilon, 
                                            beta_epsilon = beta_epsilon)
```

Now we want to choose what to do. Do we keep looking at stimulus 1 or do we look away? 

To figure this out, we need to choose proportional to expected information gain vs. the constant information gain from looking away.  

To do that, we compute EIG. 

- calculate posterior predictive distribution:
- compute p(z|theta, epsilon) --> will give P(z_ij=0) and P(z_ij=1)
- get EIG with entropy:
if true z = 1: P(z_ij=0) * (previous - alternative entropy) + P(z_ij=1) * (previous - entropy)
else if z = 0: P(z_ij=0) * (previous - entropy) + P(z_ij=1) * (previous - alternative entropy)
- get EIG with KL:
if true z = 1: P(z_ij=0) * KL(previous dist||alternative dist) + P(z_ij=1) * KL(previous dist||current dist)
else if z = 0: P(z_ij=0) * KL(previous dist||current dist) + P(z_ij=1) * KL(previous dist||alternative dist)

Let's do KL divergence for starters. So we want to compute D_{KL} for t2 under the two possible scenarios, and then weight these by the posterior predictive probability of the two scenarios. 

```{r}
o2_s1 <- obs_matrix
o2_s1[2,1] <- 0
o2_s2 <- obs_matrix
o2_s2[2,1] <- 1

posterior_o2_s1 <- grid_with_theta_and_epsilon(grid_theta = grid_theta, 
                                               grid_epsilon = grid_epsilon, 
                                               noisy_observation = head(o2_s1,2), 
                                               alpha_prior = alpha_prior, 
                                               beta_prior= beta_prior, 
                                               alpha_epsilon = alpha_epsilon, 
                                               beta_epsilon = beta_epsilon)

posterior_o2_s2 <- grid_with_theta_and_epsilon(grid_theta = grid_theta, 
                                               grid_epsilon = grid_epsilon, 
                                               noisy_observation = head(o2_s2,2), 
                                               alpha_prior = alpha_prior, 
                                               beta_prior= beta_prior, 
                                               alpha_epsilon = alpha_epsilon, 
                                               beta_epsilon = beta_epsilon)


```

Now let's think through how to compute the KL and posterior predictive.

Here's a simple version where we pretend epsilon = 0. 

```{r}
# this will have underflow issues for small numbers
dkl <- function (x,y) {
  sum(x * log(x / y)) 
}

# very simple posterior predictive
post_pred <- function(theta, p_theta, heads = TRUE) {
  ifelse(heads, 
         sum(theta * p_theta), 
         1 - sum(theta * p_theta))
}

dkl_s1 <- dkl(posterior_o2_s1$posterior, posterior_o1$posterior)
dkl_s2 <- dkl(posterior_o2_s2$posterior, posterior_o1$posterior)

post_pred_s1_0 <- post_pred(posterior_o1$theta, posterior_o1$posterior, heads = FALSE)
post_pred_s1_1 <- post_pred(posterior_o1$theta, posterior_o1$posterior, heads = TRUE)

EIG = post_pred_s1_0 * dkl_s1 + post_pred_s1_1 * dkl_s2
```

Let's make it a little more complicated and compute it when we keep track of the full posterior over epsilon and theta. 

Now we will want to compute the probability (based on theta AND epsilon) that:
y = 1, z = 1
y = 1, z = 0
y = 0, z = 1, 
y = 0, z = 0

BUT we will want to sum the probability across the values of Y because all we care about is the prediction about z!

```{r}
posterior_o1 <- 
  grid_with_theta_and_epsilon_has_epsilon(grid_theta = grid_theta, 
                                          grid_epsilon = grid_epsilon, 
                                          noisy_observation = obs_matrix[1,], 
                                          alpha_prior = alpha_prior, 
                                          beta_prior= beta_prior, 
                                          alpha_epsilon = alpha_epsilon, 
                                          beta_epsilon = beta_epsilon)

posterior_o2_s1 <- 
  grid_with_theta_and_epsilon_has_epsilon(grid_theta = grid_theta, 
                                          grid_epsilon = grid_epsilon, 
                                          noisy_observation = head(o2_s1,2), 
                                          alpha_prior = alpha_prior, 
                                          beta_prior= beta_prior, 
                                          alpha_epsilon = alpha_epsilon, 
                                          beta_epsilon = beta_epsilon)

posterior_o2_s2 <- 
  grid_with_theta_and_epsilon_has_epsilon(grid_theta = grid_theta, 
                                          grid_epsilon = grid_epsilon, 
                                          noisy_observation = head(o2_s2,2), 
                                          alpha_prior = alpha_prior, 
                                          beta_prior= beta_prior, 
                                          alpha_epsilon = alpha_epsilon, 
                                          beta_epsilon = beta_epsilon)


```

Now we have posteriors, we need to do posterior predictives. For the normal model it's:

$$\sum_{\theta} {p(y|\theta) p(\theta)}$$

But we have this annoying noisy observation thing. So it's

$$\sum_{\theta, \epsilon} {p(z | y, \epsilon) p(y | \theta) p(\theta) p(\epsilon)}$$

```{r}
# post_pred <- function(theta, p_theta, heads = TRUE) {
#    ifelse(heads, 
#           sum(theta * p_theta), 
#           1 - sum(theta * p_theta))
# }

noisy_post_pred <- function(theta, epsilon, posterior, heads = TRUE) {
  # probability of a head is:
  # (didn't get flipped * it's a head) + (got flipped + it was a tail)
  p_1 <- sum(((1 - epsilon) * theta * posterior) + 
               (epsilon * (1-theta) * posterior))
  
  # the alternative:
  # (didn't get flipped * it's a tail) + (got flipped + it's a head)
  # p_0 <- sum(((1 - epsilon) * (1 - theta) * posterior) + 
  #                 (epsilon * theta * posterior))
  
  # flip if you are asking for 0 probability
  ifelse(heads, p_1, 1 - p_1)
  
}
```

Now use these to compute the full EIG, now in the noisy scenario.  

```{r}
dkl_s1 <- dkl(posterior_o2_s1$posterior, posterior_o1$posterior)
dkl_s2 <- dkl(posterior_o2_s2$posterior, posterior_o1$posterior)

post_pred_s1_0 <- noisy_post_pred(posterior_o1$theta, 
                                  posterior_o1$epsilon,
                                  posterior_o1$posterior, heads = FALSE)
post_pred_s1_1 <- noisy_post_pred(posterior_o1$theta, 
                                  posterior_o1$epsilon,
                                  posterior_o1$posterior, heads = TRUE)

EIG = post_pred_s1_0 * dkl_s1 + post_pred_s1_1 * dkl_s2
```

# Functionalizing & sampling looking trajectories



now p-lookaway is going to be a luce choice rule of 

$$p(keep looking) = \frac{EIG^\lambda}{(EIG^\lambda + C^\lambda)}$$

Setting up stimuli and book-keeping.

```{r}
C = .005 # expected information gain from the world
max_obs <- 100 # need to pad with NAs

# generate stimuli up front
simple_stimuli <- generate_creature_sequence(block_length = trials_per_block, 
                                             deviant_positions = deviant_positions,  
                                             total_feature = num_features, 
                                             feature_theta = 0.8, 
                                             feature_number = num_features_simple, 
                                             dissimilar_ratio = dissimilarity_ratio) 


# need to track the actual observations 
observations <- list()

# book keeping data frame
df <- tibble(t = rep(NA,max_obs),
             stimulus_idx = rep(NA,max_obs), 
             dkl_if_0 = rep(NA,max_obs),
             dkl_if_1 = rep(NA,max_obs),
             post_pred = rep(NA,max_obs), 
             EIG = rep(NA,max_obs), 
             p_look_away = rep(NA,max_obs), 
             look_away = rep(NA,max_obs))

# which stimulus are we looking at
stimulus_idx <- 1
t <- 1
```

Here's the big `while` loop that samples new stimuli. 

```{r}
set.seed(42)
# now iterate through timesteps
while (stimulus_idx <= trials_per_block) {
  for (f in 1:num_features) {
    # get our observation
    observations[[t]] <- noisy_observation_creature(simple_stimuli[[stimulus_idx]],
                                                    epsilon = epsilon)    
    
    # book keeping
    df$t[t] <- t
    df$stimulus_idx[t] <- stimulus_idx
    
    # get posterior
    obs_mat <- matrix(unlist(observations), ncol = num_features, byrow = TRUE)
    
    posterior_at_t <- 
      grid_with_theta_and_epsilon_has_epsilon(grid_theta = grid_theta, 
                                              grid_epsilon = grid_epsilon, 
                                              noisy_observation = obs_mat, 
                                              alpha_prior = alpha_prior, 
                                              beta_prior= beta_prior, 
                                              alpha_epsilon = alpha_epsilon, 
                                              beta_epsilon = beta_epsilon)
    
    
    # make possible scenarios
    observations[[t+1]] <- NA
    observations[[t+1]][f] <- FALSE
    obs_mat_plus_0 <- matrix(unlist(observations), ncol = num_features, byrow = TRUE)
    observations[[t+1]] <- NA
    observations[[t+1]][f] <- TRUE
    obs_mat_plus_1 <- matrix(unlist(observations), ncol = num_features, byrow = TRUE)
    
    # update posteriors for both scenarios
    posterior_at_t_plus_1_if_0 <- 
      grid_with_theta_and_epsilon_has_epsilon(grid_theta = grid_theta, 
                                              grid_epsilon = grid_epsilon, 
                                              noisy_observation = obs_mat_plus_0, 
                                              alpha_prior = alpha_prior, 
                                              beta_prior= beta_prior, 
                                              alpha_epsilon = alpha_epsilon, 
                                              beta_epsilon = beta_epsilon)
    
    posterior_at_t_plus_1_if_1 <- 
      grid_with_theta_and_epsilon_has_epsilon(grid_theta = grid_theta, 
                                              grid_epsilon = grid_epsilon, 
                                              noisy_observation = obs_mat_plus_1, 
                                              alpha_prior = alpha_prior, 
                                              beta_prior= beta_prior, 
                                              alpha_epsilon = alpha_epsilon, 
                                              beta_epsilon = beta_epsilon)
    
    # compute EIG 
    df$dkl_if_0[t] <- dkl(posterior_at_t_plus_1_if_0$posterior, 
                          posterior_at_t$posterior)
    df$dkl_if_1[t] <- dkl(posterior_at_t_plus_1_if_1$posterior, 
                          posterior_at_t$posterior)
    
    df$post_pred[t] <- noisy_post_pred(posterior_at_t$theta, 
                                       posterior_at_t$epsilon,
                                       posterior_at_t$posterior)
    
    df$EIG[t] = (1-df$post_pred[t]) * df$dkl_if_0[t] + df$post_pred[t] * df$dkl_if_1[t] 
    
    # flip a coin with p_keep_looking weight
    df$p_look_away[t] = C / (df$EIG[t] + C)
    df$look_away[t] = rbinom(1, 1, prob = df$p_look_away[t]) == 1
    
    # if heads, learn from another sample of the same stimulus
    if (df$look_away[t]) {
      stimulus_idx <- stimulus_idx + 1
    }
    
    t <- t + 1 
  }
}
```

Now plot what happened. Here's a plot of EIGs.

```{r}
ggplot(df, aes(x = t, y = EIG, col = as_factor(stimulus_idx))) + 
  geom_point() + 
  geom_hline(yintercept = C, lty = 2) + 
  geom_vline(xintercept = df$t[df$stimulus_idx==deviant_positions][1], lty = 3) + 
  ylim(0,.1) 
```

And here's a plot of trial lengths. 

```{r}
lens <- df %>%
  group_by(stimulus_idx) %>%
  summarise(length = n()) %>%
  filter(!is.na(stimulus_idx))

ggplot(lens, aes(x = stimulus_idx, y = length)) + 
  geom_point() 

```

# Level up - multiple sims.

Hard to see what's going on here, so let's run 100 trials of this type. 

```{r}
# generate stimuli up front
simple_stimuli <- 
  generate_creature_sequence(block_length = trials_per_block, 
                             deviant_positions = deviant_positions,  
                             total_feature = num_features, 
                             feature_theta = 0.8, 
                             feature_number = num_features_simple, 
                             dissimilar_ratio = dissimilarity_ratio) 

C = .005 # expected information gain from the world
max_obs <- 100 # need to pad with NAs

# sim function
sims <- tibble(i = 1:10, 
               idx = 1:10) %>%
  split(.$i) %>%
  map_df(function(d) 
  { 
    # need to track the actual observations 
    observations <- list()
    
    # book keeping data frame
    df <- tibble(t = rep(NA,max_obs),
                 stimulus_idx = rep(NA,max_obs), 
                 dkl_if_0 = rep(NA,max_obs),
                 dkl_if_1 = rep(NA,max_obs),
                 post_pred = rep(NA,max_obs), 
                 EIG = rep(NA,max_obs), 
                 p_look_away = rep(NA,max_obs), 
                 look_away = rep(NA,max_obs))
    
    # which stimulus are we looking at
    stimulus_idx <- 1
    t <- 1
    
    # now iterate through timesteps
    while (stimulus_idx <= trials_per_block) {
      for (f in 1:num_features) {
        # get our observation
        observations[[t]] <- 
          noisy_observation_creature(simple_stimuli[[stimulus_idx]],
                                                        epsilon = epsilon)    
        
        # book keeping
        df$t[t] <- t
        df$stimulus_idx[t] <- stimulus_idx
        
        # get posterior
        obs_mat <- matrix(unlist(observations), ncol = num_features, byrow = TRUE)
        
        posterior_at_t <- 
          grid_with_theta_and_epsilon_has_epsilon(grid_theta = grid_theta, 
                                                  grid_epsilon = grid_epsilon, 
                                                  noisy_observation = obs_mat, 
                                                  alpha_prior = alpha_prior, 
                                                  beta_prior= beta_prior, 
                                                  alpha_epsilon = alpha_epsilon, 
                                                  beta_epsilon = beta_epsilon)
        
        
        # make possible scenarios
        observations[[t+1]] <- NA
        observations[[t+1]][f] <- FALSE
        obs_mat_plus_0 <- matrix(unlist(observations), ncol = num_features, byrow = TRUE)
        observations[[t+1]] <- NA
        observations[[t+1]][f] <- TRUE
        obs_mat_plus_1 <- matrix(unlist(observations), ncol = num_features, byrow = TRUE)
        
        # update posteriors for both scenarios
        posterior_at_t_plus_1_if_0 <- 
          grid_with_theta_and_epsilon_has_epsilon(grid_theta = grid_theta, 
                                                  grid_epsilon = grid_epsilon, 
                                                  noisy_observation = obs_mat_plus_0, 
                                                  alpha_prior = alpha_prior, 
                                                  beta_prior= beta_prior, 
                                                  alpha_epsilon = alpha_epsilon, 
                                                  beta_epsilon = beta_epsilon)
        
        posterior_at_t_plus_1_if_1 <- 
          grid_with_theta_and_epsilon_has_epsilon(grid_theta = grid_theta, 
                                                  grid_epsilon = grid_epsilon, 
                                                  noisy_observation = obs_mat_plus_1, 
                                                  alpha_prior = alpha_prior, 
                                                  beta_prior= beta_prior, 
                                                  alpha_epsilon = alpha_epsilon, 
                                                  beta_epsilon = beta_epsilon)
        
        # compute EIG 
        df$dkl_if_0[t] <- dkl(posterior_at_t_plus_1_if_0$posterior, 
                              posterior_at_t$posterior)
        df$dkl_if_1[t] <- dkl(posterior_at_t_plus_1_if_1$posterior, 
                              posterior_at_t$posterior)
        
        df$post_pred[t] <- noisy_post_pred(posterior_at_t$theta, 
                                           posterior_at_t$epsilon,
                                           posterior_at_t$posterior)
        
        df$EIG[t] = (1-df$post_pred[t]) * df$dkl_if_0[t] + df$post_pred[t] * df$dkl_if_1[t] 
        
        # flip a coin with p_keep_looking weight
        df$p_look_away[t] = C / (df$EIG[t] + C)
        df$look_away[t] = rbinom(1, 1, prob = df$p_look_away[t]) == 1
        
        # if heads, learn from another sample of the same stimulus
        if (df$look_away[t]) {
          stimulus_idx <- stimulus_idx + 1
        }
        
        t <- t + 1 
      }
    }
    
    df <- filter(df, !is.na(stimulus_idx))
    df$idx <- d$idx
    
    return(df)
  })
```

Now examine. 

```{r}
lengths <- sims %>%
  group_by(idx, stimulus_idx) %>%
  summarise(length = n()) 

ggplot(lengths, aes(x = stimulus_idx, y = length)) + 
  geom_jitter(width = .1, height = 0, alpha = .5) + 
  geom_smooth() + 
  geom_vline(xintercept = lengths$stimulus_idx[lengths$stimulus_idx==deviant_positions][1], lty = 3) 
```


