---
title: "dependent observation"
author: "anjie"
date: "7/22/2021"
output: html_document
---
```{r}
library(tidyverse)
library(here)
library(matrixStats)
library(profvis)

source(here("helper/get_eig_faster.R"))
source(here("helper/get_stimuli.R"))
source(here("helper/get_observation.R"))
source(here("helper/main_simulation_under_construction.r"))
source(here("helper/main_simulation_testing.r"))

source(here("helper/params_helper_experiment.r"))

num_features = 1
num_features_simple = 1
num_feature_complex = 1
trials_per_block = 6
deviant_positions = c()
feature_theta = 0.9
dissimilarity_ratio = 0.2
noise_parameter = 0.001

## grid approximation related 
grid_theta <- seq(0.1, 1, 0.2)
grid_epsilon <- seq(0.1, 1, 0.2)
alpha_prior = 1
beta_prior = 1
alpha_epsilon = 1 
beta_epsilon = 1

## eig related 
env_eig = 0.008
max_obs = 500

## experiment related 
subject_n = 10

simple_stimuli <- generate_creature_sequence(
  block_length = trials_per_block, 
  deviant_positions = deviant_positions,  # takes a vector, 
  total_feature = num_features, 
  feature_theta = feature_theta, 
  feature_number = num_features_simple, 
  dissimilar_ratio = dissimilarity_ratio)

complex_stimuli <- generate_creature_sequence(
  block_length = trials_per_block, 
  deviant_positions = deviant_positions,  # takes a vector, 
  total_feature = num_features, 
  feature_theta = feature_theta, 
  feature_number = num_feature_complex, 
  dissimilar_ratio = dissimilarity_ratio)
```




```{r}
grid_theta <- seq(0.001, 1, 0.02)
grid_epsilon <- seq(0.001, 1, 0.02)
res_flat <- main_simulation_testing(subject = 1,
                simple_stimuli %>% filter(trial_number == 1), 
                noise_parameter = 0, 
                eig_from_world =  env_eig/10,
                max_observation = max_obs, # should this be per trial or in total? currently per trial 
                grid_theta = grid_theta, 
                grid_epsilon = grid_epsilon, 
                alpha_prior = 1, 
                beta_prior = 1,
                alpha_epsilon = alpha_epsilon, 
                beta_epsilon = beta_epsilon, 
                forced_exposure = FALSE,
                forced_sample = NULL, 
                show_posterior = TRUE) 


res_false_focus <- main_simulation_testing(subject = 1,
                  simple_stimuli %>% filter(trial_number == 1), 
                noise_parameter = 0, 
                eig_from_world = env_eig/10,
                max_observation = max_obs, # should this be per trial or in total? currently per trial 
                grid_theta = grid_theta, 
                grid_epsilon = grid_epsilon, 
                alpha_prior = 1, 
                beta_prior = 3,
                alpha_epsilon = alpha_epsilon, 
                beta_epsilon = beta_epsilon, 
                forced_exposure = FALSE,
                forced_sample = NULL, 
                show_posterior = TRUE) 

res_true_focus <- main_simulation_testing(subject = 1,
                simple_stimuli %>% filter(trial_number == 1), 
                noise_parameter = 0, 
                eig_from_world =  env_eig/10,
                max_observation = max_obs, # should this be per trial or in total? currently per trial 
                grid_theta = grid_theta, 
                grid_epsilon = grid_epsilon, 
                alpha_prior = 3, 
                beta_prior = 1,
                alpha_epsilon = alpha_epsilon, 
                beta_epsilon = beta_epsilon, 
                forced_exposure = FALSE,
                forced_sample = NULL, 
                show_posterior = TRUE) 

f_res_flat <- main_simulation_testing(subject = 1,
                simple_stimuli %>% mutate(V1 = as.logical(1-V1)) %>% filter(trial_number == 1), 
                noise_parameter = 0, 
                eig_from_world =  env_eig/10,
                max_observation = max_obs, # should this be per trial or in total? currently per trial 
                grid_theta = grid_theta, 
                grid_epsilon = grid_epsilon, 
                alpha_prior = 1, 
                beta_prior = 1,
                alpha_epsilon = alpha_epsilon, 
                beta_epsilon = beta_epsilon, 
                forced_exposure = FALSE,
                forced_sample = NULL, 
                show_posterior = TRUE) 


f_res_false_focus <- main_simulation_testing(subject = 1,
                 simple_stimuli %>% mutate(V1 = as.logical(1-V1)) %>% filter(trial_number == 1), 
                noise_parameter = 0, 
                eig_from_world =  env_eig/10,
                max_observation = max_obs, # should this be per trial or in total? currently per trial 
                grid_theta = grid_theta, 
                grid_epsilon = grid_epsilon, 
                alpha_prior = 1, 
                beta_prior = 3,
                alpha_epsilon = alpha_epsilon, 
                beta_epsilon = beta_epsilon, 
                forced_exposure = FALSE,
                forced_sample = NULL, 
                show_posterior = TRUE) 

f_res_true_focus <- main_simulation_testing(subject = 1,
                 simple_stimuli %>% mutate(V1 = as.logical(1-V1)) %>% filter(trial_number == 1), 
                noise_parameter = 0, 
                eig_from_world =  env_eig/10,
                max_observation = max_obs, # should this be per trial or in total? currently per trial 
                grid_theta = grid_theta, 
                grid_epsilon = grid_epsilon, 
                alpha_prior = 3, 
                beta_prior = 1,
                alpha_epsilon = alpha_epsilon, 
                beta_epsilon = beta_epsilon, 
                forced_exposure = FALSE,
                forced_sample = NULL, 
                show_posterior = TRUE) 



plot_posterior(res_false_focus, 1, 30)
plot_posterior(f_res_false_focus, 1, 30)

plot_posterior(res_flat, 1, 30)
plot_posterior(f_res_flat, 1, 30)

plot_posterior(res_true_focus, 1, 30)
plot_posterior(f_res_true_focus, 1, 30)


plot_kl_materials(get_eig_materials(res_false_focus))
plot_kl_materials(get_eig_materials(f_res_false_focus))

plot_kl_materials(get_eig_materials(res_flat))
plot_kl_materials(get_eig_materials(f_res_flat))

plot_kl_materials(get_eig_materials(res_true_focus))
plot_kl_materials(get_eig_materials(f_res_true_focus))

```

```{r}
calculate_posteriror(res_false_focus, 10)
calculate_posteriror(f_res_false_focus, 10)
calculate_posteriror(res_flat, 10)
calculate_posteriror(f_res_flat,  10)

calculate_posteriror(res_true_focus,  10)
calculate_posteriror(f_res_true_focus, 10)

```


```{r}

bind_rows(res_flat[[1]] %>% mutate(prior = "a1b1_true"), 
          res_false_focus[[1]] %>% mutate(prior = "a1b10_true"), 
          res_true_focus[[1]] %>% mutate(prior = "a10b1_true"), 
          f_res_flat[[1]] %>%  mutate(prior = "a1b1_false"), 
          f_res_false_focus[[1]] %>% mutate(prior = "a1b10_false"), 
          f_res_true_focus[[1]] %>% mutate(prior = "a10b1_false"), )%>% 
  mutate(stimulus_idx = as.factor(stimulus_idx)) %>% 
  ggplot(aes(x = t, y = EIG, color = stimulus_idx, shape = V1)) + 
  geom_point() + 
  facet_wrap(~prior) 


res[[1]] %>% 
  mutate(stimulus_idx = as.factor(stimulus_idx)) %>% 
  ggplot(aes(x = t, y = p_look_away, color = stimulus_idx)) + 
  geom_point() + 
  facet_wrap(~subject_id)
```
```{r}
res[[1]]  %>% 
  filter(V1 == FALSE)
```

#0.01539623 0.10661700

the second dropped, and the thirteenth dropped, why? 
```{r}

# get kl if true 
get_kl_with_obs <- function(res, obs = TRUE){
  eig_materials <- res[[4]]
  kl_list = list()
  
  for (i in seq(1:length(eig_materials))){
    if (length(eig_materials[[i]]) != 0){
    
    if(obs == TRUE){
      kl_list[[i]] <- eig_materials[[i]][[1]][[1]]
    }else{
       kl_list[[i]] <- eig_materials[[i]][[1]][[2]]
    }
    }
  }
  return (kl_list)
}

get_pp_with_obs <- function(res, obs = TRUE){
  eig_materials <- res[[4]]
  pp_list = list()
  
  for (i in seq(1:length(eig_materials))){
    if (length(eig_materials[[i]]) != 0){
    
    if(obs == TRUE){
      pp_list[[i]] <- eig_materials[[i]][[2]][[1]]
    }else{
       pp_list[[i]] <- eig_materials[[i]][[2]][[2]]
    }
    }
  }
  return (pp_list)
}

get_eig_materials <- function(res){
  res[[1]]$kl_if_true <- tibble("kl_if_true" = get_kl_with_obs(res) %>% 
  unlist())

res[[1]]$kl_if_false <- tibble("kl_if_false" = get_kl_with_obs(res, FALSE) %>% 
  unlist())

res[[1]]$pp_if_true <- tibble("pp_if_true" = get_pp_with_obs(res) %>% 
  unlist())

res[[1]]$pp_if_false <- tibble("pp_if_false" = get_pp_with_obs(res, FALSE) %>% 
  unlist())


}

get_eig_materials <- function(res){
  res[[1]]$kl_if_true <- get_kl_with_obs(res) %>% 
  unlist()

res[[1]]$kl_if_false <- get_kl_with_obs(res, FALSE) %>% 
  unlist()

res[[1]]$pp_if_true <- get_pp_with_obs(res) %>% 
  unlist()

res[[1]]$pp_if_false <- get_pp_with_obs(res, FALSE) %>% 
  unlist()

return(res[[1]])

}


plot_kl_materials <- function(res_with_kl){
  plot_kl_if_true <- res_with_kl %>% 
  #filter(stimulus_idx == 1) %>% 
  ggplot(aes(x = t, y = kl_if_true, color = V1)) + 
  geom_point() + 
  theme(legend.position = "none") +
  labs(subtitle = "kl when hypothetical obs is TRUE")

plot_pp_if_true <- res_with_kl %>% 
  #filter(stimulus_idx == 1) %>% 
  ggplot(aes(x = t, y = pp_if_true, color = V1)) + 
  geom_point() + 
    theme(legend.position = "none") +
  labs(subtitle = "pp when hypothetical obs is TRUE")



plot_eig_from_both_sides <- res_with_kl %>% 
  mutate(eig_true_side = kl_if_true * pp_if_true, 
         eig_false_side = kl_if_false * pp_if_false, 
         sum_sides = eig_true_side + eig_false_side) %>% 
  pivot_longer(c(eig_true_side, eig_false_side, sum_sides), 
               names_to = "hypothetical_obs", 
               values_to = "value") %>% 
 #filter(stimulus_idx == 1) %>% 
  #filter(hypothetical_obs != "eig_true_side") %>% 
  ggplot(aes(x = t, y = value, color = hypothetical_obs)) + 
  geom_point() + 
  labs(title = "eig trajectory")

plot_kl_if_false <- res_with_kl %>% 
  #filter(stimulus_idx == 1) %>% 
  ggplot(aes(x = t, y = kl_if_false, color = V1)) + 
  geom_point() + 
    theme(legend.position = "none") +

  labs(subtitle = "kl when hypothetical obs is FALSE")

plot_pp_if_false <- res_with_kl %>% 
 # filter(stimulus_idx == 1) %>% 
  ggplot(aes(x = t, y = pp_if_false, color = V1)) + 
  geom_point() + 
    theme(legend.position = "none") +

  labs(subtitle = "pp when hypothetical obs is FALSE")

plots <- cowplot::plot_grid(plot_kl_if_true, plot_pp_if_true, 
                              plot_kl_if_false, plot_pp_if_false, 
                            ncol = 2)
cowplot::plot_grid(plots, plot_eig_from_both_sides, ncol = 1)
  
}

```

```{r}

plot_kl_materials(get_eig_materials(res_flat))
plot_kl_materials(get_eig_materials(f_res_false_focus))# all false, a1b10
plot_kl_materials(get_eig_materials(f_res_true_focus))# all false, a10b1


plot_kl_materials(get_eig_materials(res_true_focus)) # all true, a10b1 

plot_kl_materials(get_eig_materials(res_false_focus))# all true, a1b10


```

```{r}
plot_posterior <- function(res, start_time, end_time){
  for (t in seq(start_time:end_time)){
    for (f in seq(1:length(res[[2]][[t]]))){
      res[[2]][[t]][[f]]$t <- t  
    }
  }
  bind_rows(res[[2]][start_time:end_time]) %>% 
    filter(epsilon == 0.001) %>% 
    ggplot(aes(x = theta, y = posterior, color = t,group = t)) + 
    geom_point()+
    geom_line()
   
}

calculate_posteriror <- function(res, max_t){
  posteiror_for_t <- c()
  
   for (t in length(1:max_t)){
    
    posterior_at_t <- c()
    
    for (f in seq(1:length(res[[2]][[t]]))){
      res[[2]][[t]][[f]]$t <- t 
      posterior_at_t <-  c(posterior_at_t, sum(res[[2]][[t]][[f]]$posterior))
    }
    
    posteiror_for_t <- c(posteiror_for_t, sum(posterior_at_t))

   }
  return(posteiror_for_t)
  
  
  
}

plot_hypothetical_posterior <- function(res, start_time, end_time){
  
  l_hypothetical_posterior <- list()
  for (t in seq(start_time:end_time)){
    true_hypothetical_posterior <- res[[3]][[t]][[1]]
    true_hypothetical_posterior$t <- t 
    true_hypothetical_posterior$obs <- TRUE
    
    false_hypothetical_posterior <- res[[3]][[t]][[2]]
    false_hypothetical_posterior$t <- t 
    false_hypothetical_posterior$obs <- FALSE
    
    hypothetical_posterior_for_t <- bind_rows(true_hypothetical_posterior, false_hypothetical_posterior)
    l_hypothetical_posterior[[t]] <- hypothetical_posterior_for_t
  }
  
  
  bind_rows(l_hypothetical_posterior) %>% 
    #filter(epsilon == 0.001) %>% 
    ggplot(aes(x = theta, y = posterior, color = t,group = t)) + 
    geom_point()+
    geom_line() + 
    facet_wrap(~obs)
   
}

```


```{r}
plot_hypothetical_posterior(res_false_focus, 1, 10)
plot_hypothetical_posterior(res_true_focus, 1, 10)

```


```{r}
plot_posterior_all_epsilon <- function(res, time_step, max_time){
  for (t in seq(1:max_time)){
    for (f in seq(1:length(res[[2]][[t]]))){
      res[[2]][[t]][[f]]$t <- t  
    }
  }
  bind_rows(res[[2]]) %>% 
    filter(t %in% time_step) %>% 
    ggplot(aes(x = theta, y = posterior, color = epsilon)) + 
    geom_point()+
    #geom_line() + 
    facet_wrap(~t)
   
}

plot_posterior_all_epsilon(res_false_focus, c(1, 10), 30)

sum(res_false_focus[[2]][[1]][[1]]$posterior)
sum(res_false_focus[[2]][[10]][[1]]$posterior)

```


```{r}
```





why does the false side have such a big drop in kl? 
```{r}
#res[[2]][[10]]
t10 <- res[[2]][[10]]

t10_true <- res[[3]][[10]][[1]]
t10_false <- res[[3]][[10]][[2]]


t9 <- res[[2]][[9]]
t9_true <- res[[3]][[9]][[1]]
t9_false <- res[[3]][[9]][[2]]

t10_false_plot <- t10[[1]] %>% 
  mutate(type = "t10_real_posterior") %>% 
  #bind_rows(t10_true %>% mutate(type = "t10_eig_true")) %>% 
  bind_rows(t10_false %>% mutate(type = "t10_eig_false")) %>% 
  ggplot(aes(x = theta, y = posterior, color = type)) + 
  geom_point()+ 
  labs(subtitle = "after misobservation, false")

t10_true_plot <- t10[[1]] %>% 
  mutate(type = "t10_real_posterior") %>% 
  bind_rows(t10_true %>% mutate(type = "t10_eig_true")) %>% 
  #bind_rows(t10_false %>% mutate(type = "t10_eig_false")) %>% 
  ggplot(aes(x = theta, y = posterior, color = type)) + 
  geom_point()+ 
  labs(subtitle = "after misobservation, true")


t9_false_plot <- t9[[1]] %>% 
  mutate(type = "t9_real_posterior") %>% 
  #bind_rows(t10_true %>% mutate(type = "t10_eig_true")) %>% 
  bind_rows(t10_false %>% mutate(type = "t9_eig_false")) %>% 
  ggplot(aes(x = theta, y = posterior, color = type)) + 
  geom_point()+ 
  labs(subtitle = "before misobservation, false")

t9_true_plot <- t9[[1]] %>% 
  mutate(type = "t9_real_posterior") %>% 
  bind_rows(t10_true %>% mutate(type = "t9_eig_true")) %>% 
  #bind_rows(t10_false %>% mutate(type = "t10_eig_false")) %>% 
  ggplot(aes(x = theta, y = posterior, color = type)) + 
  geom_point()+ 
  labs(subtitle = "before misobservation, true")

get_kl <- function (post_posterior,prev_posterior) {
  sum(post_posterior * log(post_posterior / prev_posterior)) 
}
noisy_post_pred <- function(theta, epsilon, posterior, heads = TRUE) {
  p_1 <- sum(((1 - epsilon) * theta * posterior) + 
               (epsilon * (1-theta) * posterior))
  
  ifelse(heads, p_1, 1 - p_1)
  
}


cowplot::plot_grid(t10_false_plot, t10_true_plot, 
                   t9_false_plot, t9_true_plot, ncol = 2)

#prev_posterior 
#t10[[1]]$posterior
#post_posterior 
#t10_true$posterior 


# KL when TRUE 
print("When there's a misobservation:")
print("kl when true")
sum(t10_true$posterior * log( t10_true$posterior / t10[[1]]$posterior
))
print("pp when true")
noisy_post_pred(t10[[1]]$theta, t10[[1]]$epsilon, t10[[1]]$posterior, TRUE)

# KL when false 
print("kl when false")
sum(t10_false$posterior * log(t10_false$posterior / t10[[1]]$posterior
))
print("pp when false")
noisy_post_pred(t10[[1]]$theta, t10[[1]]$epsilon, t10[[1]]$posterior, FALSE)


print("When there's no misobservation:")
print("kl when true")
sum(t9_true$posterior * log( t9_true$posterior / t9[[1]]$posterior
))
print("pp when true")
noisy_post_pred(t9[[1]]$theta, t9[[1]]$epsilon, t9[[1]]$posterior, TRUE)

# KL when false 
print("kl when false")
sum(t9_false$posterior * log(t9_false$posterior / t9[[1]]$posterior
))
print("pp when false")
noisy_post_pred(t9[[1]]$theta, t9[[1]]$epsilon, t9[[1]]$posterior, FALSE)



```


```{r}
t10_false_plot <- t10[[1]] %>% 
  mutate(type = "t10_real_posterior") %>% 
  #bind_rows(t10_true %>% mutate(type = "t10_eig_true")) %>% 
  bind_rows(t10_false %>% mutate(type = "t10_eig_false")) %>% 
  ggplot(aes(x = epsilon, y = posterior, color = type)) + 
  geom_point()+ 
  labs(subtitle = "after misobservation, false")

t10_true_plot <- t10[[1]] %>% 
  mutate(type = "t10_real_posterior") %>% 
  bind_rows(t10_true %>% mutate(type = "t10_eig_true")) %>% 
  #bind_rows(t10_false %>% mutate(type = "t10_eig_false")) %>% 
  ggplot(aes(x = epsilon, y = posterior, color = type)) + 
  geom_point()+ 
  labs(subtitle = "after misobservation, true")


t9_false_plot <- t9[[1]] %>% 
  mutate(type = "t9_real_posterior") %>% 
  #bind_rows(t10_true %>% mutate(type = "t10_eig_true")) %>% 
  bind_rows(t10_false %>% mutate(type = "t9_eig_false")) %>% 
  ggplot(aes(x = epsilon, y = posterior, color = type)) + 
  geom_point()+ 
  labs(subtitle = "before misobservation, false")

t9_true_plot <- t9[[1]] %>% 
  mutate(type = "t9_real_posterior") %>% 
  bind_rows(t10_true %>% mutate(type = "t9_eig_true")) %>% 
  #bind_rows(t10_false %>% mutate(type = "t10_eig_false")) %>% 
  ggplot(aes(x = epsilon, y = posterior, color = type)) + 
  geom_point()+ 
  labs(subtitle = "before misobservation, true")


cowplot::plot_grid(t10_false_plot, t10_true_plot, 
                   t9_false_plot, t9_true_plot, ncol = 2)
```



```{r}
obs  %>% 
  #mutate(type = "t") %>% 
  #bind_rows(obs_f %>% mutate(type = "t+1_false")) %>%
  #bind_rows(obs_t %>% mutate(type = "t+1_true" )) %>% 
  ggplot(aes(x = theta, y = posterior, color = epsilon)) + 
  geom_point()


obs_t %>% 
  ggplot(aes(x = theta, y = posterior, color = epsilon)) + 
  geom_point()

obs_f %>% 
  ggplot(aes(x = theta, y = posterior, color = epsilon)) + 
  geom_point()
```

```{r}
# the posterior at first time point (FALSE)
obs_f <- res[[2]][[1]]
# the posterior for (FALSE TRUE)
obs_ft <- res[[3]][[1]][[1]] 
# the posterior for (FALSE FALSE)
obs_ff <- res[[3]][[1]][[2]]


```

```{r}
obs_f[[1]] %>% 
  ggplot(aes(x = theta, y = posterior, color = epsilon)) + 
  geom_point()

obs_ft %>% 
  ggplot(aes(x = theta, y = posterior, color = epsilon)) + 
  geom_point()

obs_ff %>% 
  ggplot(aes(x = theta, y = posterior, color = epsilon)) + 
  geom_point()
```

















# THE TEST
the posterior produced for EIG at t should be the same with one of the real observation 
```{r}
subject = 1
stimuli_sequence = simple_stimuli
noise_parameter = noise_parameter 
eig_from_world = env_eig
max_observation = max_obs # should this be per trial or in total? currently per trial 
grid_theta = seq(0.01, 1, 0.3) 
grid_epsilon = seq(0.01, 1, 0.3) 
alpha_prior = 1 
beta_prior = 1
alpha_epsilon = 1 
beta_epsilon = 1 
```


```{r}

  feature_number <- ncol(stimuli_sequence[startsWith(names(stimuli_sequence), 
                                                     "V")])
  total_trial_number = max(stimuli_sequence$trial_number)
  
  # df for keeping track of model behavior, 
  df_model <-  initialize_model(subject, eig_from_world, max_observation)
  m_observation <- initialize_m_observation(feature_number, max_observation, stimuli_sequence)
  ll_df_posterior <- initialize_ll_df_posterior(grid_theta, grid_epsilon, max_observation, feature_number)
  ll_df_z_given_theta <- initialize_ll_df_z_given_theta(grid_theta, grid_epsilon, max_observation, feature_number)
  
  
  
  

  
  # material for calculating df_posterior 
  df_lp_theta_epsilon <- get_df_lp_theta_epsilon(grid_theta, grid_epsilon, 
                                                 alpha_prior, beta_prior, 
                                                 alpha_epsilon, beta_epsilon)
  df_lp_y_given_theta = tibble(
    "theta" = grid_theta, 
    "lp_y_ONE_given_theta" =  lp_yi_given_theta(yi = 1, theta = grid_theta ), 
    "lp_y_ZERO_given_theta" = lp_yi_given_theta(yi = 0, theta = grid_theta )
  )

  
  stimulus_idx <- 1
  t <- 1
  
    
for(t in seq(1, 2)){
 
  print(t)
   df_model$t[[t]] = t
    df_model$stimulus_idx[[t]] = stimulus_idx
  
    
    #get stimulus
    current_stimulus <- stimuli_sequence[stimulus_idx,]
    
    #get observation 
    # SET RANDOM SEED TO SEE IF IT IS RIGHT

    current_observation <- noisy_observation_creature(
      creature = current_stimulus[,str_detect(names(current_stimulus), "V")], 
      n_sample = 1, 
      epsilon = noise_parameter
    )
  
    # add to current observation 
    m_observation[t, ] <- current_observation
  
    
   
     ## need to update posterior df here 
    
    current_unique_possible_combinations <- get_unique_combination(t, 
                                                                m_observation, 
                                                                feature_number)
  
    feature_pos <- current_unique_possible_combinations$feature_pos
    
    all_possible_combinations <- expand_grid(
      current_unique_possible_combinations, 
      hypothetical_observation = c(TRUE, FALSE)
      
    ) 
    
    n_possible_combination <- nrow(all_possible_combinations)
    all_possible_combinations$kl <- rep(NA_real_, n_possible_combination)
    all_possible_combinations$post_predictives <- rep(NA_real_, n_possible_combination)
    
    feature_occurence <- na.omit(as.vector(sapply(feature_pos, function(x){first(na.omit(x))})))
    for (index in feature_occurence){
      
      # this needs to be changed! we need to take into account of whether sth is taken from a new stimulus or not
    
      
      ll_df_z_given_theta[[t]][[index]] <- get_df_lp_z_given_theta(t, 
                                                                   df_lp_y_given_theta,
                                                                   ll_df_z_given_theta, 
                                                                   stimulus_idx,   # needs to be about each observation, not each stimulus  
                                                                   index, 
                                                                   df_model, 
                                                                   m_observation,
                                                                   current_observation, 
                                                                   grid_theta, grid_epsilon, 
                                                                   alpha_prior, beta_prior)
      
      
      
     
        unnormalized_log_posterior <- ll_df_z_given_theta[[t]][[index]]$lp_z_given_theta + 
          df_lp_theta_epsilon$lp_theta + 
          df_lp_theta_epsilon$lp_epsilon

      
      ll_df_posterior[[t]][[index]]$unnormalized_log_posterior <- unnormalized_log_posterior
      ll_df_posterior[[t]][[index]]$log_posterior <-  ll_df_posterior[[t]][[index]]$unnormalized_log_posterior - matrixStats::logSumExp( ll_df_posterior[[t]][[index]]$unnormalized_log_posterior)
      ll_df_posterior[[t]][[index]]$posterior <- exp(ll_df_posterior[[t]][[index]]$log_posterior)
      
      print("real occurence")
      print("real observation:")
      print(m_observation %>% janitor::remove_empty())
     print(ll_df_posterior[[t]][[index]]$posterior)
    }
    
    for (i in 1:feature_number){
      
      # find corresponding calculated value 
      calculated_value_index <- match(TRUE, sapply(feature_pos, function(x){i %in% x}))
      calculated_value_index_in_ll <- feature_occurence[[calculated_value_index]]
      ll_df_posterior[[t]][[i]] <- ll_df_posterior[[t]][[calculated_value_index_in_ll]]
      ll_df_z_given_theta[[t]][[i]] <- ll_df_z_given_theta[[t]][[calculated_value_index_in_ll]]
    }
    
    
    prev_posterior_list <- ll_df_posterior[[t]][feature_occurence]
    prev_z_given_theta_list <- ll_df_z_given_theta[[t]][feature_occurence]
    
    post_posterior_list <- lapply(seq(1, n_possible_combination),
                                  function(x){
                                    expand_grid(theta = grid_theta, 
                                                epsilon = grid_epsilon)
                                  })
    
    last_t_for_last_stimulus = ifelse(stimulus_idx == 1, 1,
                                      max((df_model[df_model$stimulus_idx == stimulus_idx-1,])$t, na.rm = TRUE)
    )
    prev_z_given_theta_last_stimulus <- ll_df_z_given_theta[[last_t_for_last_stimulus]][feature_occurence]
    
    
    for (i in 1:n_possible_combination){
      
      
      all_hypothetical_observations_on_this_stimulus = c((unlist(all_possible_combinations$unique_combination[[i]]))[last_t_for_last_stimulus:t], 
                             all_possible_combinations$hypothetical_observation[[i]])
      
      
      post_posterior_df = post_posterior_list[[i]]
      
      prev_observation_posterior = prev_posterior_list[[ceiling(i/2)]]
      prev_observation_z_given_theta = prev_z_given_theta_list[[ceiling(i/2)]]
      prev_last_stimulus_observation_z_given_theta = prev_z_given_theta_last_stimulus[[ceiling(i/2)]]
      
      print("in eig calculation")
      post_df_z_given_theta = eig_get_df_lp_z_given_theta(t,
                                                       df_model,
                                                       prev_last_stimulus_observation_z_given_theta,
                                                       all_hypothetical_observations_on_this_stimulus, # contains unique combination + hypothetical scenarios 
                                                       grid_theta, grid_epsilon, 
                                                       df_lp_y_given_theta)
      print("hypothetical observation:")
      print(all_hypothetical_observations_on_this_stimulus)

      
      post_posterior_list[[i]]$unnormalized_log_posterior <- post_df_z_given_theta$lp_z_given_theta +  df_lp_theta_epsilon$lp_theta + 
        df_lp_theta_epsilon$lp_epsilon
      
      post_posterior_list[[i]]$log_posterior <-   post_posterior_list[[i]]$unnormalized_log_posterior - matrixStats::logSumExp(post_posterior_list[[i]]$unnormalized_log_posterior)
      post_posterior_list[[i]]$posterior <- exp(post_posterior_list[[i]]$log_posterior)
      print(post_posterior_list[[i]]$posterior)
    
    }
    
    
     for (s in 1:n_possible_combination){
      
      all_possible_combinations$kl[s] <- get_kl(post_posterior_list[[s]]$posterior, 
                                                prev_posterior_list[[ceiling(s/2)]]$posterior)
      all_possible_combinations$post_predictives[s] <- noisy_post_pred(prev_posterior_list[[ceiling(s/2)]]$theta, 
                                                                       prev_posterior_list[[ceiling(s/2)]]$epsilon, 
                                                                       prev_posterior_list[[ceiling(s/2)]]$posterior, 
                                                                       all_possible_combinations$hypothetical_observation[s]) 
      
      
    }
    print("post preditives")
    print(all_possible_combinations$post_predictives)
    
    t = t + 1 
    stimulus_idx <- stimulus_idx + 1
}
```



```{r}
all_hypothetical_observations_on_this_stimulus
post_posterior_list
```




















```{r}
sim_res %>% 
  group_by(subject_id, stimulus_idx) %>% 
  mutate()

sim_res %>% 
  get_sim_res
```

```{r}
sim_res %>% 
  ggplot(aes(x = t, y = EIG, color = stimulus_idx)) + 
  geom_point()  + 
  facet_wrap(~subject_id)
```



```{r}
all_false <- simple_stimuli %>% 
  mutate(V1 = rep(FALSE, 6))


mix_test_stim_posterior_ll <- main_simulation_uc(subject = 1,
                all_false, 
                noise_parameter = noise_parameter, 
                eig_from_world = env_eig ,
                max_observation = max_obs, # should this be per trial or in total? currently per trial 
                grid_theta = grid_theta, 
                grid_epsilon = grid_epsilon, 
                alpha_prior = 1, 
                beta_prior = 1,
                alpha_epsilon = alpha_epsilon, 
                beta_epsilon = beta_epsilon, 
                forced_exposure = FALSE,
                forced_sample = NULL, 
                show_posterior = TRUE) 

for (t in seq(1: length(mix_test_stim_posterior_ll))){
  for (f in seq(1:length(mix_test_stim_posterior_ll[[t]]))){
    mix_test_stim_posterior_ll[[t]][[f]]$t <- t  
  }
}

bind_rows(mix_test_stim_posterior_ll) %>% 
  filter(!is.na(posterior)) %>% 
  #filter(t%%2 == 0) %>% 
  ggplot(aes(x  = theta, y = posterior)) + 
  geom_point() + 
  facet_wrap(~t)
```



```{r}
env_eig <- 0.001
simple_sim <- main_simulations(subject_n = 50,
                simple_stimuli, 
                noise_parameter = noise_parameter, 
                eig_from_world = env_eig,
                max_observation = max_obs, # should this be per trial or in total? currently per trial 
                grid_theta = grid_theta, 
                grid_epsilon = grid_epsilon, 
                alpha_prior = 1, 
                beta_prior = 1,
                alpha_epsilon = alpha_epsilon, 
                beta_epsilon = beta_epsilon, 
                forced_exposure = FALSE,
                forced_sample = NULL) %>% 
  mutate(type = "simple_stimuli")

complex_sim <- main_simulations(subject_n = 50,
                complex_stimuli, 
                noise_parameter = noise_parameter, 
                eig_from_world = env_eig,
                max_observation = max_obs, # should this be per trial or in total? currently per trial 
                grid_theta = grid_theta, 
                grid_epsilon = grid_epsilon, 
                alpha_prior = 1, 
                beta_prior = 12,
                alpha_epsilon = alpha_epsilon, 
                beta_epsilon = beta_epsilon, 
                forced_exposure = FALSE,
                forced_sample = NULL) %>% 
  mutate(type = "complex_stimuli")


```



```{r}
cm_sim <- bind_rows(simple_sim %>% get_sim_res() %>% mutate(complexity = "complex"), complex_sim %>% get_sim_res() %>% mutate(complexity = "simple"))

cm_eig <- bind_rows(simple_sim, complex_sim) %>% 
  group_by(subject_id, type, stimulus_idx) %>% 
  summarise(total_eig = sum(EIG))


cm_sim %>% 
  ggplot(aes(x = stimulus_idx, y = sample_n, color = complexity)) + 
  geom_jitter(alpha = .1) + 
  stat_summary(fun.data = "mean_cl_boot") + 
  ylim(0, 50) 

bind_rows(simple_sim, complex_sim) %>% 
   ggplot(aes(x = t, y = p_look_away, color = type, alpha = .3))  + 
  stat_summary(fun.data = "mean_cl_boot")  

cm_sim %>% 
  left_join(cm_eig, by = c("subject_id", "stimulus_idx")) %>% 
  ggplot(aes(x = total_eig, y = sample_n, color = type)) + 
  geom_point() + 
  facet_wrap(~type) + 
  ylim(0, 100)
cm_eig


bind_rows(simple_sim, complex_sim) %>% 
  ggplot(aes(x = t, y = EIG, color = type, alpha = .3))  + 
  stat_summary(fun.data = "mean_cl_boot") 
 # facet_wrap(~type)
```






i'm going to systematically vary: 

- forced sample 
- noise parameter 
- eig_envrs 
- alpha_prior 
- beta_prior 

```{r}


subject_n <- 40
stimuli_sequence <- test_background
noise_parameter <- 0.001
eig_from_world <- seq(0.001, 0.02, 0.002)
max_observation <- 100

alpha_prior <- 5
beta_prior <- 4
forced_sample = seq(1, 30, 2)

background_params <- expand_grid(
  subject_n,
  #stimuli_sequence = nest(test_background), 
  noise_parameter,
  eig_from_world, 
  max_observation,
  alpha_prior, 
  beta_prior, 
  forced_sample
) %>%
  mutate(
    sim_id = row_number(),
    type = "background"
  )

deviant_params <- expand_grid(
  subject_n,
  #timuli_sequence = nest(test_deviant), 
  noise_parameter,
  eig_from_world, 
  max_observation,
  alpha_prior, 
  beta_prior, 
  forced_sample
) %>%
  mutate(
    sim_id = row_number(),
    type = "deviant"
  )
  

params <- bind_rows(background_params, deviant_params) 

params
params %>% data.frame() 
  
test_background
test_deviant
params
```


with the same parameter i'm going to feed them the two stimuli
the end table i wants (each row is a column)

- forced_sample_n 
- noise_parameter 
- eig_envrs 
- alpha_prior 
- beta_prior 
- mean_sample_n_background 
- mean_sample_n_deviant
- sd_sample_n_background 
- sd_sample_n_deviant 




```{r}


simulation_wrappter <- function(subject_n, noise_parameter, 
                                eig_from_world, max_observation, alpha_prior, beta_prior, forced_sample, sim_id, 
                                type){
  
  b_sim_df <- main_simulations(subject_n, 
                             test_background, 
                             noise_parameter, 
                             eig_from_world,
                             max_observation, # should this be per trial or in total? currently per trial 
                             grid_theta = grid_theta, 
                             grid_epsilon = grid_epsilon, 
                             alpha_prior, 
                             beta_prior,
                             alpha_epsilon = alpha_epsilon, 
                             beta_epsilon = beta_epsilon, 
                             forced_exposure = TRUE,
                             forced_sample) %>% 
    get_sim_res() %>% 
    get_sim_eva() 

    b_sim_df$sim_id <- sim_id
    b_sim_df$type <- "background"
    
     d_sim_df <- main_simulations(subject_n, 
                             test_deviant, 
                             noise_parameter, 
                             eig_from_world,
                             max_observation, # should this be per trial or in total? currently per trial 
                             grid_theta = grid_theta, 
                             grid_epsilon = grid_epsilon, 
                             alpha_prior, 
                             beta_prior,
                             alpha_epsilon = alpha_epsilon, 
                             beta_epsilon = beta_epsilon, 
                             forced_exposure = TRUE,
                             forced_sample) %>% 
    get_sim_res() %>% 
    get_sim_eva() 
     
    d_sim_df$sim_id <- sim_id
    d_sim_df$type <- "deviant"
      
    print(sim_id)
    return(bind_rows(b_sim_df, d_sim_df))  
  }


noise_sim_res_redux <- pmap_dfr(params, 
     simulation_wrappter
     )

saveRDS(noise_sim_res_redux, file = "noise_sim_res.rds")

#saveRDS(sim_res, file = "sim_res.rds")
```

```{r}
noise_sim_res
```
```{r}
forced_exposure_df <- noise_sim_res_redux %>% 
  filter(stimulus_idx == 1) %>% 
  distinct(n, sim_id, mean_lt, stimulus_idx) %>% 
  pivot_wider(names_from = stimulus_idx, 
              values_from = mean_lt) %>%
  rename(forced_exposure_n = "1") %>% 
  select(sim_id, forced_exposure_n)
 

fam_df <- noise_sim_res_redux %>% 
  filter(stimulus_idx == 2) %>% 
  left_join(forced_exposure_df, by = "sim_id") %>% 
  mutate(ci.upper = mean_lt + (1.96 * sd_lt)/sqrt(n), 
         ci.lower = mean_lt - (1.96 * sd_lt)/sqrt(n)) 
  
fam_df %>% 
  ggplot(aes(x = forced_exposure_n, y = mean_lt, color = type)) + 
  geom_jitter(alpha = .5) + 
  geom_smooth(method = "lm") + 
  theme_classic()
```

```{r}

params_df <- params %>% 
  mutate(temp_id = paste0(sim_id, type)) %>% 
  select(forced_sample,
      noise_parameter, 
         eig_from_world, 
         alpha_prior, 
         beta_prior, 
         temp_id) %>% 
  left_join(noise_sim_res_redux %>%  filter(stimulus_idx == 2) %>% mutate(temp_id = paste0(sim_id, type)), 
            by = "temp_id") %>% 
  select(-temp_id)
```

```{r}
params_df %>% 
  ggplot(aes(x = mean_lt, fill = type)) + 
  geom_density(alpha = .3) + 
  facet_wrap(~forced_sample)
```



```{r}
params_df %>% 
  mutate(ci.upper = mean_lt + (1.96 * sd_lt)/sqrt(n), 
         ci.lower = mean_lt - (1.96 * sd_lt)/sqrt(n)) %>% 
  ggplot() + 
  geom_pointrange(aes(x = forced_sample, y = mean_lt, ymax = ci.upper, ymin = ci.lower, 
                      color = type), position = position_jitter(width = .5)) + 
  facet_wrap(~eig_from_world)


params_df %>% 
  mutate(ci.upper = mean_lt + (1.96 * sd_lt)/sqrt(n), 
         ci.lower = mean_lt - (1.96 * sd_lt)/sqrt(n)) %>% 
  filter(eig_from_world == 0.001) %>% 
  ggplot() + 
  geom_pointrange(aes(x = forced_sample, y = mean_lt, ymax = ci.upper, ymin = ci.lower, 
                      color = type), position = position_jitter(width = .5)) + 
  facet_wrap(~sim_id)
  
```



```{r}
params_df %>% 
  mutate(prior_print = paste("a", alpha_prior, "b", beta_prior, sep = "_")) %>% 
  ggplot(aes(x = forced_exposure_n, y = mean_lt, color = type)) + 
  geom_jitter(alpha = .5) + 
  geom_smooth(method = "lm") + 
  theme_classic()+ 
  facet_wrap(~forced_exposure_n)
```


```{r}
bckgrd_df <- params_df %>% 
  filter(type == "background") %>% 
 pivot_wider(names_from = type, 
             values_from = mean_lt) %>% 
  select(sim_id, background, forced_exposure_n)
  
deviant_df <- params_df %>% 
  filter(type == "deviant") %>% 
 pivot_wider(names_from = type, 
             values_from = mean_lt) %>% 
  select(sim_id, deviant)


sim_res_df <- bckgrd_df %>% 
  left_join(deviant_df, by = "sim_id") 


sim_res_df %>% 
  ggplot(aes(x =bckgrd_minus_deviant, fill = forced_exposure_n)) + 
  geom_density(alpha = .1)  +
  geom_vline(xintercept = 0) + 
  theme_classic() + 
  facet_wrap(~as.numeric(forced_exposure_n))

```






```{r}
saveRDS(params_df, file = "noise_sim_res.rds")
```






























