---
title: "params visualization"
author: "anjie & gal"
date: "4/30/2022"
output:
  html_document:
    code_folding: hide
---

```{r}
knitr::opts_chunk$set(message=FALSE, warning = FALSE)
```


```{r}
library(here)
library(tidyverse)
```

unless otherwise specified, the parameters used in this markdown are: 

- grid_mu_theta <- seq(-1, 1, .2)
- grid_sig_sq = seq(0.001, 2, 0.2)
- grid_y <- seq(-1, 1, .2)
- grid_epsilon = seq(0.001, 1, 0.2)
- hypothetical_obs_grid_n = 3 
- mu_prior = 0
- V_prior = 0.001
- alpha_prior = 1
- beta_prior = 1
- mu_epsilon = 0.001 (prior for noise)
- sd_epsilon = 4 (prior for noise)


- epsilon = 0.000001
- world_EIGs = 0.0001 
(and note for demo purpose we are currently not using decision noise)

background stimulus: 0.1
novel stimulus: 0.8


# stability across priors 

## for mu 

for non-extreme values, the patterns are pretty robust. but not so when the priors are out of the grid range. 

```{r}
prior_comparison <- readRDS(here("param_exploration/cached_res/B_0.1_d_0.8_mu_comparison.RDS"))
 model_res <- prior_comparison %>% 
    ungroup() %>% 
    mutate(res_row_number = row_number()) %>% 
    filter(res_row_number %% 2 == 0) %>% 
    unnest(results)

 model_res %>%   
    group_by(sequence_scheme, mu_prior, params_id, stimulus_idx) %>% 
    filter(!is.na(stimulus_idx)) %>% 
    summarise(sample_n = n()) %>% 
    ggplot(aes(x = stimulus_idx, y = sample_n)) + 
    stat_summary(position = position_dodge(width = .1),fun.data = "mean_cl_boot") + 
    facet_grid(mu_prior~sequence_scheme)
```

## for sig sq

```{r}
prior_sig_comparison <- readRDS(here("param_exploration/cached_res/B_0.1_D_0.8_sig_comparison.RDS"))
 model_res <- prior_sig_comparison %>% 
    ungroup() %>% 
    mutate(res_row_number = row_number()) %>% 
    filter(res_row_number %% 2 == 0) %>% 
    unnest(results)

 model_res %>%   
    group_by(sequence_scheme, V_prior, params_id, stimulus_idx) %>% 
    filter(!is.na(stimulus_idx)) %>% 
    summarise(sample_n = n()) %>% 
    ggplot(aes(x = stimulus_idx, y = sample_n)) + 
    stat_summary(position = position_dodge(width = .1),fun.data = "mean_cl_boot") + 
    facet_grid(V_prior~sequence_scheme)
```

# Influence of grid / stimulus value 

when the deviant is greater than the grid range, familiarity preference? or just non-interpretable? 


```{r}
b1d2 <- readRDS(here("param_exploration/cached_res/B_0.1_D_0.2.RDS")) %>% mutate(deviant = 0.2)
b1d8 <- readRDS(here("param_exploration/cached_res/B_0.1_D_0.8.RDS"))%>% mutate(deviant = 0.8)
b1d12 <- readRDS(here("param_exploration/cached_res/B_0.1_D_1.2.RDS")) %>% mutate(deviant = 1.2)

model_res <- bind_rows(b1d2, b1d8, b1d12) %>% 
   ungroup() %>% 
    mutate(res_row_number = row_number()) %>% 
    filter(res_row_number %% 2 == 0) %>% 
    unnest(results)

model_res %>%   
    group_by(sequence_scheme, stimulus_idx, deviant) %>% 
    filter(!is.na(stimulus_idx)) %>% 
    summarise(sample_n = n()) %>% 
    ggplot(aes(x = stimulus_idx, y = sample_n, color = as.factor(deviant), group = deviant)) + 
    geom_point(position = position_dodge(width = .3))+ 
    geom_line(position = position_dodge(width = .3)) + 
     #stat_summary(position = position_dodge(width = .1),fun.data = "mean_cl_boot") + 
    facet_wrap(~sequence_scheme)
```


also interestingly note the asymmetry (probably related to prior)

```{r}
b8d1 <- readRDS(here("param_exploration/cached_res/B_0.8_D_0.1.RDS")) %>% mutate(stimulus = "b_0.8_d_0.1")
b1d8 <- readRDS(here("param_exploration/cached_res/B_0.1_D_0.8.RDS"))%>% mutate(stimulus =  "b_0.1_d_0.8")

model_res <- bind_rows(b8d1, b1d8) %>% 
   ungroup() %>% 
    mutate(res_row_number = row_number()) %>% 
    filter(res_row_number %% 2 == 0) %>% 
    unnest(results)

model_res %>%   
    group_by(sequence_scheme, stimulus_idx, stimulus) %>% 
    filter(!is.na(stimulus_idx)) %>% 
    summarise(sample_n = n()) %>% 
    ggplot(aes(x = stimulus_idx, y = sample_n, color = as.factor(stimulus), group = stimulus)) + 
    geom_point(position = position_dodge(width = .3))+ 
    geom_line(position = position_dodge(width = .3)) + 
     #stat_summary(position = position_dodge(width = .1),fun.data = "mean_cl_boot") + 
    facet_wrap(~sequence_scheme)
```

EIG VIZ

```{r}
b1d8_EIG <- b8d1$results[[2]]

 b1d8_EIG %>% ggplot(aes(x = t, y = EIG, color = as.factor(stimulus_idx))) + 
  stat_summary(position = position_dodge(width = .1),fun.data = "mean_cl_boot")
```


