---
title: "understanding patterns of sim"
author: "anjie"
date: "9/27/2021"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(matrixStats)


source(here("helper/make_scheme_and_params.r"))
source(here("helper/main_simulation_testing.r"))

source(here("helper/get_eig_faster.R"))
source(here("helper/get_stimuli.R"))
source(here("helper/get_observation.R"))
source(here("helper/main_simulation_under_construction.r"))
source(here("helper/params_helper_experiment.r"))
```

```{r}
grid_theta <- seq(0.1, 1, 0.2)
grid_epsilon <- seq(0.1, 1, 0.2)

n = 100
alpha_epsilon = c(1)
beta_epsilon = c(1)
alpha_priors = c(1)
beta_priors = c(1)
noise_parameters = c(0.03)
eigs_from_world = c(0.001)
max_observation = 500
fixed_length_complexity = TRUE 
feature_space_n = 10
simple_feature_n = 1
complex_feature_n = 5
dissimilar_ratio = 1
sequence_scheme = c("BBBBBD")
complexity = c("complex", "simple")

full_params_df <- make_simulation_params(n, 
                                   sequence_scheme, 
                                   complexity, 
                                   alpha_priors, 
                                   beta_priors, 
                                   alpha_epsilon, 
                                   beta_epsilon, 
                                   noise_parameter, 
                                   eig_from_world, 
                                   max_observation, 
                                   fixed_length_complexity, 
                                   feature_space_n, 
                                   simple_feature_n, 
                                   complex_feature_n, 
                                   dissimilar_ratio) 


```




```{r}
S <- (full_params_df %>% 
  filter(complexity == "complex") %>% 
  select(stimuli_sequence) %>% 
  pull())[[1]][[1]] %>% 
  select(V2, V3) %>% 
  rename(V1 = V2, V2 = V3) %>% 
  mutate(trial_number = row_number()) %>% 
  select(V2, trial_number) %>% 
  rename(V1 = V2)
S
```

if (show_posterior){
    res_list <- list()
    
    obs_df <- as.data.frame(m_observation) %>% mutate(t = row_number())
    df_model <- df_model %>% 
      left_join(obs_df, by = "t")
    res_list[[1]] <- df_model %>% filter(!is.na(EIG))
    res_list[[2]] <- ll_df_posterior
    res_list[[3]] <- ll_df_hypothetical_posterior
    res_list[[4]] <- ll_eig_materials
    names(res_list) <- c("df_model", "real posterior", "hypothetical_posterior", "eig_materials")
    return (res_list)
    
    
  }
  
   l_eig_materials[[1]] <- l_kl
  l_eig_materials[[2]] <- l_pp
  
```{r}
a1b50_res[[1]]
```


```{r}
a1b50_res <- main_simulation_testing(subject = 1,
                               stimuli_sequence = super_S, 
                               noise_parameter = 0, 
                               eig_from_world = 0.001,
                               max_observation = 50, # should this be per trial or in total? currently per trial 
                               grid_theta = grid_theta, 
                               grid_epsilon = grid_epsilon, 
                               alpha_prior = 1, 
                               beta_prior = 50,
                               alpha_epsilon = 1, 
                               beta_epsilon = 1, 
                               forced_exposure = FALSE,
                               forced_sample = 5, 
                               show_posterior = TRUE)
```

```{r}
 S
super_S <- bind_rows(S %>% filter(trial_number != 6), 
                     S %>% filter(trial_number != 6),
                     S %>% filter(trial_number != 6), 
                     S %>% filter(trial_number != 6), 
                     S %>% filter(trial_number != 6), 
                     S) %>% 
  mutate(trial_number = row_number())
super_S
```

noisy_post_pred <- function(theta, epsilon, posterior, heads = TRUE) {
  p_1 <- sum(((1 - epsilon) * theta * posterior) + 
               (epsilon * (1-theta) * posterior))
  
  ifelse(heads, p_1, 1 - p_1)
  
}



```{r}



a50b1_res[[2]][[2]][[1]] %>% 
  ggplot(aes(x = theta, y = posterior, color = epsilon)) + 
  geom_point()
a1b50_res[[2]][[2]][[1]]%>% 
  ggplot(aes(x = theta, y = posterior, color = epsilon)) + 
  geom_point()

noisy_post_pred(a50b1_res[[2]][[2]][[1]]$theta, a50b1_res[[2]][[2]][[1]]$epsilon, a50b1_res[[2]][[2]][[1]]$posterior, TRUE)
noisy_post_pred(a1b50_res[[2]][[2]][[1]]$theta, a1b50_res[[2]][[2]][[1]]$epsilon, a1b50_res[[2]][[2]][[1]]$posterior, TRUE)

noisy_post_pred(a50b1_res[[2]][[1]]$theta, a50b1_res[[2]][[1]]$epsilon, a50b1_res[[2]][[1]]$posterior, FALSE)
noisy_post_pred(a1b50_res[[2]][[1]]$theta, a1b50_res[[2]][[1]]$epsilon, a1b50_res[[2]][[1]]$posterior, FALSE)

noisy_post_pred(a50b1_res[[3]][[1]]$theta, a50b1_res[[3]][[1]]$epsilon, a50b1_res[[3]][[1]]$posterior, TRUE)
noisy_post_pred(a1b50_res[[3]][[1]]$theta, a1b50_res[[3]][[1]]$epsilon, a1b50_res[[3]][[1]]$posterior, TRUE)
```


```{r}
a50b1_res <- main_simulation_testing(subject = 1,
                               stimuli_sequence = super_S, 
                               noise_parameter = 0, 
                               eig_from_world = 0.001,
                               max_observation = 50, # should this be per trial or in total? currently per trial 
                               grid_theta = grid_theta, 
                               grid_epsilon = grid_epsilon, 
                               alpha_prior = 50, 
                               beta_prior = 1,
                               alpha_epsilon = 1, 
                               beta_epsilon = 1, 
                               forced_exposure = FALSE,
                               forced_sample = 5, 
                               show_posterior = TRUE)
```

```{r}
a1b1_res <- main_simulation_testing(subject = 1,
                               stimuli_sequence = super_S, 
                               noise_parameter = 0, 
                               eig_from_world = 0.001,
                               max_observation = 50, # should this be per trial or in total? currently per trial 
                               grid_theta = grid_theta, 
                               grid_epsilon = grid_epsilon, 
                               alpha_prior = 1, 
                               beta_prior = 1,
                               alpha_epsilon = 1, 
                               beta_epsilon = 1, 
                               forced_exposure = FALSE,
                               forced_sample = 5, 
                               show_posterior = TRUE)
```

```{r}
a1b3_res <- main_simulation_testing(subject = 1,
                               stimuli_sequence = S, 
                               noise_parameter = 0, 
                               eig_from_world = 0.001,
                               max_observation = 50, # should this be per trial or in total? currently per trial 
                               grid_theta = grid_theta, 
                               grid_epsilon = grid_epsilon, 
                               alpha_prior = 1, 
                               beta_prior = 3,
                               alpha_epsilon = 1, 
                               beta_epsilon = 1, 
                               forced_exposure = FALSE,
                               forced_sample = 5, 
                               show_posterior = TRUE)

a3b1_res <- main_simulation_testing(subject = 1,
                               stimuli_sequence = S, 
                               noise_parameter = 0, 
                               eig_from_world = 0.001,
                               max_observation = 50, # should this be per trial or in total? currently per trial 
                               grid_theta = grid_theta, 
                               grid_epsilon = grid_epsilon, 
                               alpha_prior = 3, 
                               beta_prior = 1,
                               alpha_epsilon = 1, 
                               beta_epsilon = 1, 
                               forced_exposure = FALSE,
                               forced_sample = 5, 
                               show_posterior = TRUE)
```



  
```{r}
S
plot_kl_materials(get_eig_materials(a1b50_res))
plot_kl_materials(get_eig_materials(a50b1_res))
plot_kl_materials(get_eig_materials(a1b1_res))
plot_kl_materials(get_eig_materials(a1b3_res))
plot_kl_materials(get_eig_materials(a3b1_res))

```

```{r}


```


```{r}
plot_posterior_all_epsilon(a1b3_res, seq(1, 50), 50)
plot_posterior_all_epsilon(a3b1_res, seq(1, 50), 50)

plot_posterior_all_epsilon(a1b50_res, seq(1, 50), 50)
plot_posterior_all_epsilon(a50b1_res, seq(1, 50), 50)

```



```{r}
a1b50_res[[4]][[1]]
```

